manuscript_NS_MRD_atlas

### 2020-01-22 manuscript_NS_MRD.Rmd
updated

#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006) # only uses the baseline samples. This is n=1006

@@

# reorder
# NOTE THERE IS NO EASTERN EUROPE
c.mnd.nationality <- fct_collapse(c.mnd.atlas, LOW = c("Scandinavia") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE") )

@@

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE") )

@@

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa")  
  save_plot(paste("PERMANOVA_",title_text,".png", sep = "" ), p)
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

@@

@@

## PERMANOVA significance test for group-level differences
Evaluate whether the MND grouping has a significant effect on overall gut microbiota composition

>>>>

#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:
data(atlas1006)
atlas.pseq <- baseline(atlas1006) # only uses the baseline samples. This is n=1006
atlas.pseq <- subset_samples(atlas.pseq, !is.na(mnd))

@@

# reorder
# NOTE THERE IS NO EASTERN EUROPE
c.mnd.nationality <- fct_collapse(c.mnd.atlas, LOW = c("Scandinavia") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE" ), HIGH = c("US") )

@@

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope" ,"UKIE"), HIGH = c("US") )

@@

# This calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(x,y) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 12)  + annotate("text", x = 2.5, y = 0.01, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa") + scale_fill_discrete(name = "Taxa") + theme(axis.text.y = element_text(face="italic"))
  save_plot(paste("PERMANOVA_",title_text,".png", sep = "" ), p)
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
 }

@@

save_plot(paste("akkermansia",title_text,".png", sep = "" ), p)
}

plotAtlas = plot_abundance.ak(atlas.pseq, "atlas")
plotCur = plot_abundance.ak(pseq.subset, "curated")
grid.arrange(nrow = 2, plotAtlas, plotCur)

@@

### 2020-01-22 manuscript_NS_MRD_atlas.Rmd
updated

---
title: "curatedMetagenomicData"
author: "NSharma"
date: "26/09/2019"
output:
  word_document: default
  html_document: default
---

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')

@@

eh = ExperimentHub()
myquery = query(eh, "curatedMetagenomicData")

# Get the stool samples
esl <- curatedMetagenomicData("*metaphlan_bugs_list.stool*", dryrun = FALSE)

# merge
eset <- mergeData(esl)

# the relab=FALSE is essential as it Absolute Raw Count Data
pseq = ExpressionSet2phyloseq(eset,relab=FALSE ,simplify = TRUE, phylogenetictree = TRUE)
orig_pseq <- pseq

@@

## Incidence of MND
```{r}

#THIS IS NO LONGER NEEDED.

# This block uses the raw data to examin  the incidence of MND to ensure the classificaiton is correct. It assumes that the data is in the correct place so the setwd at the top meeds to be accurate. 
# setwd('~/globalmicrobiome') 
# # For MND  nrows = 196 as thwew are other things after 196
# mnd.i.15_49 <- as.data.frame(read.csv("./rawdata/MND_I_15_49.csv", header=TRUE, nrows = 196))

# mnd.i.50_69  <- as.data.frame(read.csv("./rawdata/MND_I_50_69.csv", header=TRUE, nrows = 196))
# mnd.i.70 <- as.data.frame(read.csv("./rawdata/MND_I_70plus.csv", header=TRUE, nrows = 196))

# mnd.i.70.test <- mnd.i.70
# # Run function and change name
# mnd.i.15_49 <- mb.gather(mnd.i.15_49)
# colnames(mnd.i.15_49)[2] <- "15-49"

# # Run function and change name
# mnd.i.50_69 <- mb.gather(mnd.i.50_69)
# colnames(mnd.i.50_69)[2] <- "50-69"

# # Run function and change name
# mnd.i.70 <- mb.gather(mnd.i.70)
# colnames(mnd.i.70)[2] <- "70"

# # Join them together.   
# mnd.i <- list( mnd.i.15_49,mnd.i.50_69,mnd.i.70 ) %>% reduce(left_join, by = "Location")

# a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE","Scandinavia" ))

# mnd.i %>%
#   kable(col.names = c("","Incidence","CI","Incidence","CI","Incidence","CI")) %>%
#   kable_styling(bootstrap_options = c("striped", "hover"),full_width = F,position = "center") %>%
# add_header_above(c("Location" = 1, "15-49" = 2, "50-69" = 2, "70+" = 2)) %>%
#   add_header_above(c(" ", "Age (years)" = 6)) %>%
#   pack_rows("High MND group", 5,6)%>%
#   pack_rows("Low MND group", 2, 2) %>%
#   pack_rows("Medium MND group",1,3,4)

# # Now do the same for the curated data set
# mnd.ic.15_49 <- mb.cgather(mnd.i.15_49)
# colnames(mnd.ic.15_49)[2] <- "15-49"

# # Run function and change name
# mnd.ic.50_69 <- mb.cgather(mnd.i.50_69)
# colnames(mnd.ic.50_69)[2] <- "50-69"

# # Run function and change name
# mnd.ic.70 <- mb.cgather(mnd.i.70)
# colnames(mnd.ic.70)[2] <- "70"

# # Join them together.

# mnd.ic <- list( mnd.ic.15_49,mnd.ic.50_69,mnd.ic.70 ) %>% reduce(left_join, by = "Location")

# mnd.ic <- dplyr::select(mnd.ic, "Location",  "Value.x", "Lower.bound.x", "Upper.bound.x", "Value.y", "Lower.bound.y", "Upper.bound.y", "Value", "Lower.bound", "Upper.bound" )

# # Prep the data for the table

# mnd.ic %>%
#   kable(col.names = c("","Incidence","Upper","Lower","Incidence","Upper","Lower","Incidence","Upper","Lower"),digits=3) %>%
#   kable_styling(bootstrap_options = c("striped", "hover"),full_width = F,position = "left") %>%
# add_header_above(c("Location" = 1, "15-49" = 3, "50-69" = 3, "70+" = 3)) %>%
#   add_header_above(c(" ", "Age (years)" = 9)) %>%
#   pack_rows("Group 1", 4, 7) %>%
#   pack_rows("Group 1", 4, 7) %>%
#   pack_rows("Group 2", 8, 10)

@@

# THIS FOR THE CURATED DATA SET

# creates a grouping varaible for BMI
n <- length(sample_data(pseq)$BMI)
data <- data.frame(id=1:n)
data$bmi <- sample_data(pseq)$BMI # creates a new variable with BMI in it
# Classification based upon the ATlas data
data$bmi[data$bmi <18.5] <- "underweight"
data$bmi[data$bmi >=45] <- "super obese"
data$bmi[data$bmi >= 18.5 & data$bmi <25] <- "lean"
data$bmi[data$bmi >= 25 & data$bmi <30] <- "overweight"
data$bmi[data$bmi >= 30 & data$bmi <35] <- "obese"
data$bmi[data$bmi >= 35 & data$bmi <40] <- "severe obese"
data$bmi[data$bmi >= 40 & data$bmi <45] <- "morbid obese"
data$bmi <- factor(data$bmi)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$bmi_group = data$bmi
# Check that it works
get_variable(pseq, "bmi_group")

# Create mnd variable
#pseq <- orig_pseq
# grabs the nationality from phyloseq
c.mnd.country <- get_variable(pseq, "country" )
table(c.mnd.country)

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
# Not used BGD Bangladesh , CAN Canada, CHN, China. DNK denmark, FJI Figi, HUN Hungry, ISL isle of man , MDG, MNG mongolia, PER Iran, RUS RUSSIA,

#Adjsuted to match exactly the ATLAS data set

#central Europe = Belgium, Demark, Netherlands, Germany
#southern Europe = France, Italy, Serbia, Spain
#UKIE
#Scandinavia = Norway, Sweden Finland
#Eastern Europe = Poland

c.mnd.atlas <- fct_collapse(c.mnd.country, Scandinavia = c("NOR","SWE","FIN") ,CentralEurope = c("DNK","DEU","NLD"), SouthEurope = c("ITA","ESP","FRA"), US = "USA", UKIE = "GBR" )

#drops unused levels
c.mnd.atlas <- droplevels(c.mnd.atlas,c("ISL","HUN","EST", "SVK", "LUX", "CAN", "AUT", "BGD","FJI","CHN","MDG","MNG","PER","RUS","TZA") )

sample_data(pseq)$nationality = c.mnd.atlas

#mnd.country <- fct_collapse(mnd.country, LOW = c("NOR","SWE","FIN","DNK","LUX") , MEDIUM = c("ITA","AUT","DEU","ESP","FRA","NLD"), HIGH = c("GBR","USA","CAN") )

#mnd.country <- fct_collapse(mnd.country, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
# NOTE THERE IS NO EASTERN EUROPE
c.mnd.nationality <- fct_collapse(c.mnd.atlas, LOW = c("Scandinavia") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE") )

c.mnd.nationality <- factor(c.mnd.nationality, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(c.mnd.nationality)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$mnd = c.mnd.nationality

# checks that it has worked.
get_variable(pseq, "mnd")
table(get_variable(pseq, "mnd"))

###############################################

# THIS IS FOR THE ATLAS 1006 
# Create mnd variable
# grabs the nationality from phyloseq
a.mnd.nationality <- get_variable(atlas.pseq, "nationality" )

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope" ), HIGH = c("US","UKIE") )

@@

mb.gather <- function (x) {
    # Creates an empty data frame
    mb.group <- data.frame("Location" = c("Central Europe","Eastern Europe","Scandinavia", "Southern Europe","UKIE","USA"), "Mean" = c(1:6)) 
     # groups the countries to match the microbiome classification 
  c.europe <<- filter(x,Location=="Belgium"|Location=="Denmark"|Location=="Germany"|Location=="The Netherlands")
    e.europe <- filter(x,Location=="Poland")
    scan <- filter(x,Location=="Finland"|Location=="Norway"|Location=="Sweden")
    s.europe <- filter(x,Location=="France"|Location=="Italy"|Location=="Serbia"|Location=="Spain")
    ukie <- filter(x,Location=="United Kingdom"|Location=="Ireland")
    usa <- filter(x,Location=="United States")
    # gmeans the data and put it into the data frame

    mb.group[1,2] <- round(mean(c.europe$Value),digits = 2)
    mb.group[2,2] <- round(mean(e.europe$Value),digits = 2)
    mb.group[3,2] <- round(mean(scan$Value),digits = 2)
    mb.group[4,2] <- round(mean(s.europe$Value),digits = 2)
    mb.group[5,2] <- round(mean(ukie$Value),digits = 2)
    mb.group[6,2] <- round(mean(usa$Value),digits = 2)

    mb.group[1,3] <- paste(round(mean(c.europe$Upper.bound),digits = 2), round(mean(c.europe$Lower.bound),digits = 2), sep = "-",collapse = "-")
    mb.group[2,3] <- paste(round(mean(e.europe$Upper.bound),digits = 2), round(mean(e.europe$Lower.bound),digits = 2), sep = "-",collapse = "-")
    mb.group[3,3] <- paste(round(mean(scan$Upper.bound),digits = 2), round(mean(scan$Lower.bound),digits = 2), sep = "-",collapse = "-")
    mb.group[4,3] <- paste(round(mean(s.europe$Upper.bound),digits = 2), round(mean(s.europe$Lower.bound),digits = 2), sep = "-",collapse = "-")
    mb.group[5,3] <- paste(round(mean(ukie$Upper.bound),digits = 2), round(mean(ukie$Lower.bound),digits = 2), sep = "-",collapse = "-")
    mb.group[6,3] <- paste(round(mean(usa$Upper.bound),digits = 2), round(mean(usa$Lower.bound),digits = 2), sep = "-",collapse = "-")
    # mb.group[2,3] <- paste(e.europe$Upper.bound,"-", e.europe$Lower.bound)
    # mb.group[3,3] <- paste(scan$Upper.bound,"-", scan$Lower.bound)
    # mb.group[4,3] <- paste(s.europe$Upper.bound,"-",s.europe$Lower.bound)
    # mb.group[5,3] <- paste(ukie$Upper.bound,"-", ukie$Lower.bound )
    # mb.group[6,3] <- paste(usa$Upper.bound,"-", usa$Lower.bound)
    return(mb.group)
    }

mb.cgather <- function (x) {
    # Creates an empty data frame
    mb.group <- data.frame("Location" = c("Norway","Sweden","Finland","Denmark","Luxembourg", "Italy","Austria","Germany","Spain","France","Netherlands","United Kingdom","United States","Canada"), "Mean" = c(1:14)) 
     # groups the countries to match the microbiome classification 
  mb.group <- filter(x,Location=="Norway"|Location=="Sweden"|Location=="Finland"|Location=="Denmark"|Location=="Luxembourg"|Location=="Italy"|Location=="Austria"|Location=="Germany"|Location=="Spain"|Location=="France"|Location=="Netherlands"|Location=="United Kingdom"|Location=="United States"|Location=="Canada")

    return(mb.group)
}

my_cur_table <- function(pseq.subset){
    n <- length(sample_data(pseq.subset)$mnd)
    data <- data.frame(id=1:n)
    data$mnd <- sample_data(pseq.subset)$mnd
    data$bmi <- sample_data(pseq.subset)$BMI
    data$bmi_group <- sample_data(pseq.subset)$bmi_group
    data$age <- sample_data(pseq.subset)$age
    data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
    data$smoker <- sample_data(pseq.subset)$smoker
    data$pregnant <- sample_data(pseq.subset)$pregnant
    data$alcohol <- sample_data(pseq.subset)$alcohol
    data$age_category <- sample_data(pseq.subset)$age_category
    table1(~ age  +  bmi + bmi_group + smoker + alcohol + age_category | mnd, data=data)
}

@@

keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
pseq.f <- subset_taxa(pseq, keepotu)

pseq.c <- core(pseq, detection = 0.1/100, prevalence = 50/100)

@@

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" )

setwd("~/globalmicrobiome")

# These are useful is filtering. 
#sample_variables(pseq.subset)
#get_variable(pseq.subset, sample_variables(pseq.subset)[13])
my_cur_table(pseq.subset)
my_cur_table(pseq.c)

@@

alpha_d(atlas.pseq,"final_RLWT_2")

beta_d(pseq.subset,"final_2")

pre_perma_d(pseq.subset) 
perma_d(p.low,p.high,"RLWT_Curatefinal_highVlow")

pre_perma_d(pseq.subset) 
perma_d(p.med,p.high,"RLWT_Curatefinal_highVmed")

@@

## BETA DIVERSITY
Beta-diversity: Measures for differences between samples from different groups to identify if there are differences in the overall community composition and structure.
WE want to: quantify community divergence within a given sample set
Divergence of a given sample can be quantified as the average dissimilarity of each sample from the group mean; the dissimilarity can be quantified by beta diversity.

Divergence of a given sample set can be quantified as the average dissimilarity of each sample from the group mean; the dissimilarity can be quantified by beta diversity, for instance. This was applied in group-level comparisons for instance in Salonen et al. ISME J 2014 (they focused on homogeneity using inverse correlation, whereas here we focus on divergence using correlation but the measure is essentially the same).

 The low group tends to have smaller values, indicating that the samples are more similar to the group mean, and is less heterogeneous (has smaller spread / is more homogeneous):

## PERMANOVA significance test for group-level differences
Evaluate whether the MND grouping has a significant effect on overall gut microbiota composition

NOTHING BELOW THIS WORKS!!!!!!!!

## MICROBIOME COMPOSITION

```{r Phylum #}

# Show available ranks in the dataset
rank_names(pseq)

# Create table, number of features for each phyla
table(tax_table(pseq)[, "Phylum"], exclude = NULL)

@@

This shows a few phyla for which only one feature was observed. Those may be worth filtering.

guide_italics
ggtitle("Relative abundance") +  
```{r}

clos.ph <- aggregate_taxa(clostridium, "Phylum")

clos.ph.rel <- microbiome::transform(clos.ph, "compositional")

plot.comp.rel <- plot_composition(clos.ph.rel, x.label = "sample") + 
  theme(legend.position = "bottom", ) + theme_bw() + 
  theme(axis.text.x=element_blank()) + 
  theme(legend.title = element_text(size=18), legend.text = element_text( face="italic")) + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend(title="Phylum"))

print(plot.comp.rel)

ggsave("phylum_relative_abundance_clos.png")

plot.comp.rel + scale_fill_brewer(palette = "Paired") + theme(legend.position="none") + guide_italics
ggsave("phylum_relative_abundance_clos2.png")

@@

```{r CLostridium Bar Plot}

clos <- aggregate_taxa(clostridium, "Phylum")

pseq.cl.rel <- microbiome::transform(clos, "compositional")

plot.cl.rel <- plot_composition(pseq.cl.rel, average_by = "mnd", x.label = "sample", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

plot.cl.rel + scale_fill_brewer(palette = "Paired")
ggsave("cl_abundance_bar.png")

```{r Clostridium genus bar plots}

pseq.cl.rel <- microbiome::transform(clostridium, "compositional")

plot.cl.rel <- plot_composition(pseq.cl.rel, average_by = "mnd", x.label = "sample", plot.type = "barplot")

plot.cl.rel + scale_fill_viridis(discrete = TRUE, option = "A")

```{r clostridium heatmap}

pseq.c.ph <- aggregate_taxa(clostridium, "Phylum")

pseq.c.ph.rel <- microbiome::transform(pseq.c.ph, "compositional")

plot.c.comp.rel <- plot_composition(pseq.c.ph.rel, x.label = "sample", plot.type = "heatmap")

p <- plot.c.comp.rel + coord_flip()
p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
ggsave("clos_histogram.png")

## Community composition plotting
Classic bar plots of bacterial phyla present per sample can be useful for communicating "high level" results. These are relatively easy to interpret when major shifts in microbial communities are present, such as in this study where antibiotics are used However, they are not effective at detecting subtle shifts in communities or taxa and do not convey any statistical significance and can be subjectively interpreted. Interpretation of these plots should always be subject to subsequent statistical analysis.

```{r}

clos <- aggregate_taxa(clostridium, "Phylum")

# agglomerate at phylum level
clos_phy <- tax_glom(clos, taxrank = "Phylum")

# Transform to rel. abundance (or use ps0.ra)
clos_phyl <- transform_sample_counts(clos_phy, function(x) {x/sum(x)})

# Melt to long format for easy ggploting
clos_phylum <- psmelt(clos_phyl)

# Plot - Phylum
p.ra.phylum <- ggplot(clos_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(title = "Abundant Phylum of Clostridium(> 1%)")

p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
plot(p)
ggsave("phylum_abun_mnd_clos.png")

```{r}

get_taxa_unique(pseq, "Phylum")

```{r}

get_taxa_unique(pseq, "Genus")

```{r}

get_taxa_unique(clostridium, "Genus")

```{r}

df <- psmelt(pseq)
kable(head(df))

```{r}

head(abundances(pseq)["Faecalibacterium prausnitzii et rel.",])

>>>>

---
title: "manuscript_NS_MRD_atlas"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')

@@

@@

@@

# Create mnd variable
# grabs the nationality from phyloseq
a.mnd.nationality <- get_variable(atlas.pseq, "nationality" )

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
a.mnd.nationality <- fct_collapse(a.mnd.nationality, LOW = c("EasternEurope","Scandinavia") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE" ), HIGH = c("US") )

@@

@@

#keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
#pseq.f <- subset_taxa(pseq, keepotu)

#pseq.c <- core(pseq, detection = 0.1/100, prevalence = 50/100)

@@

#pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" )

#setwd("~/globalmicrobiome")

# These are useful is filtering. 
#sample_variables(pseq.subset)
#get_variable(pseq.subset, sample_variables(pseq.subset)[13])

@@

@@

```{r function list}

my_atlas_table(atlas.pseq)

alpha_d(atlas.pseq, "title")

alpha_d <- function(pseq.subset,title_text) {
  # gets the meta data 
  pseq.meta <- meta(pseq.subset)
  # generates the diversites
  tab <- microbiome::alpha(pseq.subset, index = "shannon")

  print("Third graph has completed")
  save_plot(paste(title_text,".png", sep = ""),p3,base_height = 11.69 )
  ggsave("/Users/maryam/Alpha_diversity_curate.png")
}

@@

```{r permanova test}

#pre_perma_d <- function(pseq.subset) { 
  p.low <<- subset_samples(atlas.subset, mnd == "LOW")
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(atlas.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(atlas.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

#}

# THis calcualtes the group PERMANOVA perma_d(p.low,p.high,"TESTHigh_vesus_Low")
#perma_d <- function(x,y,title_text) { # x & y must be either p.low,p.med or p.high
  # Convert to compositional data perma_d(p.low,p.high,"final_highVlow")
  pseqLvH <- merge_phyloseq(p.low, p.high) # pseqLvH <- merge_phyloseq(p.low,p.high)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])  
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(12, 30, 2, 1))
  #KEEP THIS  TO ENSURE THERE IS NOT A MISTAKE IN THE REORDERING BELOW
  #temp <- TeX("Top Genus")
  #png(paste("PERMOVA_",title_text,".png", sep = "" ), width = 450, height = 210,  units     = "mm",
  #res       = 1500)
  #barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
  #dev.off() 
  temp <- as.data.frame(top.coef)
  temp$label <- rownames(temp) 
  temp$label <- sub("s_*", "", temp[,2])
  #temp$colour <- ifelse(temp$top.coef < 0, "firebrick1","steelblue")
  p <<- ggplot(temp,aes(reorder(label, top.coef),top.coef,fill = top.coef > 0)) + geom_col(show.legend = FALSE,) + coord_flip() +  xlab('Taxa') + ylab('p value')  + theme_minimal(base_size = 20)  + annotate("text", x = 2.5, y = 0.02, label = "Harmful Taxa") + annotate("text", x = 19, y = -0.01, label = "Beneficial Taxa") + guide_italics
   save_plot(paste("PERMANOVA_",title_text,".png", sep = "" ), p)
   plot(p)
  #ggsave("test_permanova.png", device = "png",width = 297, height = 210, units = "mm")
# }

@@
